---
params:
  model: Arima(order=c(1, 0, 0))
  series: '2hrs ph3'
  startday: -10
  traindays: 7
  testdays: 3
title: '`r params$model` on `r params$series`'
author: 'Barbu Paul - Gheorghe'
#date: "October 17, 2018"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r includes, warning=FALSE, include=FALSE}
library(forecast)
library(ggplot2)
library(tictoc)

source('datasets.R')
source('funcs.R')
```

# Training & test data
The data has been previously cleaned, negative values were made 0.
Time series 80/20 rule goes in sequential order, not random.

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
data <- get_days(datasets[[params$series]]$series, -10, 7, 3)
autoplot(data$train, series="Train") +
  autolayer(data$test, series="Test") +
  xlab("Observation [days]") +
  ylab("Energy [Watts]") +
  ggtitle("Original data") + 
  guides(colour=guide_legend(title="Data Series"))
```

Total data points: **`r length(datasets[[params$series]]$series)`** representing **`r round(length(datasets[[params$series]]$series)/frequency(data$train))`** days.

Number of data points per day: **`r frequency(data$train)`** (gathered once every **`r (24*60)/frequency(data$train)`** minutes)

Starting from: **`r params$startday`**.

Train days: **`r params$traindays`**.

Test days: **`r params$testdays`**.

```{r subseries-plots, echo=FALSE, message=FALSE}
season.labels = seq(frequency(data$train))
ggseasonplot(data$train, season.labels = season.labels) +
  xlab("Season") +
  ylab("Energy [Watts]") +
  ggtitle('Seasonal plot for training data')

ggsubseriesplot(data$train, labels = season.labels) +
  xlab("Season") +
  ylab("Energy [Watts]") +
  ggtitle('Subseries plot for each season in the training data') +
  scale_x_continuous(breaks = 0.5 + (1:frequency(data$train)),
                     labels = 1:frequency(data$train))
```

# ACF & PACF of training data

```{r acf-pacf, echo=FALSE}
data$train %>% ggtsdisplay(main='Training Data', xlab='Observation [days]', ylab='Energy [Watts]')
```

# Box-Cox transformations

```{r boxcox, echo=FALSE}
lambda <- BoxCox.lambda(data$train)
```
Suggested BoxCox transformation on the training data: **`r lambda`**

```{r boxcox-applied, echo=FALSE}
autoplot(BoxCox(data$train, lambda)) +
  xlab("Observation [days]") +
  ylab("Energy [Watts]") +
  ggtitle("Transformed Training Data")
```

## ACF & PACF after differencing and transforming

Suggested number of differences to make data stationary: **`r ndiffs(data$train)`**.

Suggested number of seasonal differences to make data stationary: **`r nsdiffs(data$train)`**.

```{r diffs}
data$train %>% diff() %>%
  ggtsdisplay(main='First difference',
              xlab='Observation [days]',
              ylab='Energy Change [Watts]')

data$train %>% diff() %>% diff(lag=frequency(data$train)) %>%
  ggtsdisplay(main='Second difference', 
              xlab='Observation [days]', 
              ylab='Energy Acceleration [Watts]')
```

# Model fitting & evaluation

```{r model-fitting, include=FALSE}
tic("Model fitting")

#last_week %>% auto.arima(xreg=fourier(last_week, K=2), seasonal=F) -> fit
eval(parse(text=paste('data$train %>%', params$model))) -> fit

times <- toc()
```

The model **`r params$model`** was evaluated on **`r params$train`** days in **`r times$toc - times$tic`** seconds.

## Model summary

```{r summary-fit}
summary(fit)
```

## Residuals statistics

```{r include=FALSE}
fit %>% residuals() -> data$residuals
```


```{r residuals-stats, echo=FALSE}
checkresiduals(fit,
               xlab='Observation [days]',
               ylab='Energy [Watts]') 
```

Box-Ljung test:

* H~0~ states that the data points are independently distributed.
* High p-value (> 0.05) doesn't reject H~0~.
* If Q\* is large: the autocorrelations do not come from a white noise series (so Q\* should be small)

```{r residuals-pacf, echo=FALSE}
ggPacf(data$residuals) +
  ggtitle('PACF of residuals')
```

```{r qq-plot, echo=FALSE, message=FALSE}
ggplot(data.frame(y=data$residuals), aes(sample=y)) +
  stat_qq() +
  stat_qq_line() +
  ggtitle('Residuals QQ-plot')
```

## Forecasting
```{r forecast, echo=FALSE, include=FALSE}
tic('Forecast')

fit %>% forecast(h=params$testdays * frequency(data$test)) -> fcast
#fit %>% forecast(xreg=fourier(data$, K=2)) %>% autoplot()

fcast.times <- toc()
```

Time elapapsed for forecasting **`r params$testdays`** days: **`r fcast.times$toc - fcast.times$tic`** seconds.

### Accuracy of the forecasts against the test data
```{r accuracy, echo=FALSE}
accuracy(fcast, data$test)
```

### Fitted values and forecasts plot against the real data
```{r fit-forecast-plot, echo=FALSE}
fcast %>% autoplot(series='Forecast') +
  autolayer(fitted(fit), series='Fitted') + 
  autolayer(data$train, series = 'Train') +
  autolayer(data$test, series='Test') +
  xlab("Observation [days]") +
  ylab("Energy [Watts]") +
  guides(colour=guide_legend(title="Data Series"), 
         fill=guide_legend(title="Prediction Interval"))
```

# Future work

* Add other statistical tests for stationarity, differencing, etc.
* Allow the model to be ran on the manually transformed data (through a param)
* Parallel processing for evaluating multiple models at the same time

# References
* [Forecasting: Principles and Practice, Rob J Hyndman and George Athanasopoulos, Monash University, Australia](https://otexts.org/fpp2/)
* [Statistical forecasting: notes on regression and time series analysis, Robert Nau, Fuqua School of Business, Duke University](http://people.duke.edu/~rnau/411home.htm)
* [Time Series Forecasting: Creating a seasonal ARIMA model using Python and Statsmodel, Sean Abu](http://www.seanabu.com/2016/03/22/time-series-seasonal-ARIMA-model-in-python/)
* [How to Create an ARIMA Model for Time Series Forecasting in Python, Jason Brownlee](https://machinelearningmastery.com/arima-for-time-series-forecasting-with-python/)
* [ Introduction to Forecasting with ARIMA in R, Ruslana Dalinina](https://www.datascience.com/blog/introduction-to-forecasting-with-arima-in-r-learn-data-science-tutorials)
* [Seasonal periods, Rob J Hyndman](https://robjhyndman.com/hyndsight/seasonal-periods/)